\textbf{Beasts of Prey or Rational Animals?\\ Private Governance in Brazil's \textit{Jogo do Bicho}}

### 2018-10-15 paper-bicho/bicho.tex
Update 2.3

\textit{Banqueiros} solve problems of internal cooperation by providing club goods \citep{buchanan1965economic} while simultaneously shunning cheaters through selective punishments \citep{bo2005cooperation, roth1978equilibrium}. The first club good offered to \textit{bicheiros} by their bosses is private security. As the game is illegal, street sellers cannot rely on official institutions to protect themselves. Thus, the game bankers have built an extensive network of gunmen and bribed police officers to protect their employees from other criminals \citetext{\citealp[48]{chinelli1993vazio}; \citealp[51]{labronici2012paratodos}}.

@@

\newpage
\bibliography{bibliography}
\bibliographystyle{apalike}
\end{document}

>>>>

\textit{Banqueiros} solve problems of internal cooperation by providing club goods \citep{buchanan1965economic} while simultaneously shunning cheaters through selective punishments \citep{bo2005cooperation, roth1978equilibrium}. The first club good offered to \textit{bicheiros} by their bosses is private security. As the game is illegal, street sellers cannot rely on official institutions to protect themselves. Thus, the game bankers have built an extensive network of gunmen and bribed police officers to protect their employees from other criminals \citetext{\citealp[48]{chinelli1993vazio}; \citealp[51]{labronici2012paratodos}}. `Zé' (Little Joe), a bicheiro interviewed by \citet[52]{labronici2012paratodos}, described eloquently the deterring effect of the \emph{jogo do bicho} informal security personnel:

\begin{quote}
 [\dots] bums are scared and they don't mess around with us; they think there's a guard nearby or something like that. Look at all this money here! [shows the interviewer a handful of cash] It's not ours [referring to street-corner bookmakers]. And if it's not ours, it's someone else's. When I worked in Penha (\emph{a low middle-class neighbourhood in the city of Rio de Janeiro -- translator's note}), the owner of a pub close to where I used to work always asked me to stay at the front door of his pub. People know that bums are afraid of \emph{bicheiros}.
\end{quote}

But the \textit{banqueiros} did not use violence only against other criminals. They often employ violent methods against competitors and their own staff, too. \citet{jupiara2015poroes} argue that Ailton Guimarães Jorge, a former Army officer, tortured and murdered rival lottery bosses in the late 1970s. One of his former allies, Army Colonel Paulo Malhães, told the Rio de Janeiro State Truth Commission how Guimarães consolidated his power: `he went on a rampage' \citep{belem2015guimaraes}. $Castor d$e Andrade, Rio's most influential animal game \textit{banqueiro}, also employed similar methods to run his business. Andrade kept an armed bodyguard of 23 men and allegedly murdered a number of competitors. In a famous case, Andrade shot Euclides Ponar, an old \textit{banqueiro} known as `Grey-Headed Chinese',\footnote{In Portuguese, \textit{China Cabe\c{c}a Branca}.} after Ponar denounced a fraud in lottery draws in 1976. Andrade directly planned the assassination of at least other four competitors in the 1980s \citep{globo2017castor}.

@@

\newpage
\bibliography{bibliography.bib}
\bibliographystyle{apalike}
\end{document}
